Near and Mid-infrared properties of known $z\geq5$ Quasars

### 2019-06-12 tex/VHzQ.tex
Kris Kross
anoter wee update
Corrected text/ added a little more

\title[Infrared properties Very high-$z$ Quasars]
  {Near and Mid-infrared properties of known $z\geq5$ Quasars}
\author[Ross \& Cross]
{Nicholas P. Ross\thanks{E-mail: npross@roe.ac.uk} and Nicholas J. G. Cross\\ 
Institute for Astronomy, University of Edinburgh, Royal Observatory, Edinburgh, EH9 3HJ, United Kingdom\\ }

@@

Using the extended datasets described in Section~\ref{sec:NIR_data},
we select objects with at least 8 measurements observed with intervals
of at least 30 days, in at least one filter. The the average
calibrated flux over all epochs must be $>0$ (i.e. {\tt aperJky3}$>0$)
with the signal-to-noise is required to be $\geq8$. (i.e. {\tt
aperJky3/aperJky3Err}$\geq$8). The clipped median and median absolute
deviation is then calculated
\begin{equation}
      var = 1.48 \times  {\rm m.a.d.} / \bar{\epsilon}
\end{equation}
where $var$ is the index of variation, m.a.d. is the median absolute
deivation, and $\bar{\epsilon}$ is the mean of the error in each point
in the light curve, divided by the total number of points.  With these
criteria, there are 21 WFCAM and 12 VISTA objects with 1 object (VIK
J1148+0056) in common to both, and this sample of 32 is our starting
point for variability investigations.

@@

\begin{figure}
  \begin{subfigure}{}
    \centering
    \includegraphics[width=8.5cm]{../light_curves/MMTJ0215-0529LC_20190214.png}
    \caption{The infrared light curve for MMT J0215-0529 with data from 
      WFCAM (smaller solid cirlces), VIRCAM (hexagons) and NEOWISE-R (larger solid cirlces).}
    \label{fig:MMTJ0215-0529}
  \end{subfigure} 
  \begin{subfigure}{}
    \centering
    \includegraphics[width=8.5cm]{../light_curves/CFHQSJ0216-0455LC_20190214.png}
    \caption{The same as Fig.~\ref{fig:MMTJ0215-0529} except for CFHQS J0216-0455.}
    \label{fig:CFHQSJ0216-0455}
  \end{subfigure}
\end{figure}

@@

Even with well sampled data across the 3000-4000 day observed
time-scales that the WFCAM and VISTA surveys span, the $(1+z)$
time-dilation dramatically affects the sampled rest-frame timescales
sampled. Indeed, when sharp changes in a accretion rate are expected
on the system's dynamical timescale of several kiloyears
\citep[e.g.,][]{Regan2019}, then seeing any variability signature is
not expected. However, noting how in lower redshift quasars, which
also have massive $>10^{8}$ M$_{\odot}$ black holes, dramatic changes
in both continuum and line emission {\it is seen} on much shorter
timescales, continued monitoring of these objects is warrented.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%  S E C T I O  N   7         S E C T I O  N   7           S E C T I O  N   7       S E C T I O  N   7
%%  S E C T I O  N   7         S E C T I O  N   7           S E C T I O  N   7       S E C T I O  N   7
%%  S E C T I O  N   7         S E C T I O  N   7           S E C T I O  N   7       S E C T I O  N   7
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}\label{sec:conclusions}
In this study, we have, for the first time, compiled the list of all
$z>5$ spectroscopically confirmed quasars. We have assemble the NIR
($y/Y, J, H, K/K_{s}$) and MIR (WISE W1/2/3/4) photometry for these
objects, given their detection rates and SEDs. We find that:

\begin{itemize}
    \item SDSS and Pan-STARRS1 together identified over half of the VH$z$Q sample;  
    \item There remains a quasar ``redshift desert'' at $z\approx5.3-5.7$, though 
      efforts are being made to address this \citep[e.g.,][]{YangJ2018a};
    \item 97.0\% of the VH$z$Q sample is detected in one or more NIR ($ZYJHK/K_{s}$) band; 
    \item The 14 objects that are not detected in the NIR is due to lack of coverage rather than lack of depth; 
   \item 362 ( 78.2\%) VH$z$Qs are detected by WISE, e.g. in the deeper unWISE W1 catalog. 
 \item All of the $z\geq7$ quasars are detected in both unWISE W1 and W2. 
\end{itemize}

@@

@@

%\newpage

@@

\section{Near-Infrared WFCAM Science Archive SQL queries}\label{sec:SQL}
Here we give the receipe and SQL that returned the near-infrared photometry 
for the VH$z$Qs from the  WFCAM Science Archive.

The data are on the WFCAM Science Archive: \href{wsa.roe.ac.uk}{\tt wsa.roe.ac.uk}. 
Access the User Login form \href{WFCAM Science Archive}{\tt wsa.roe.ac.uk/login.html} 
with these credentials::
\begin{itemize}
    \item Username: {\tt WSERV1000} 
    \item password: {\tt highzqso} 
    \item community: {\tt nonsurvey}
\end{itemize}
Then going to the
\href{http://wsa.roe.ac.uk:8080/wsa/SQL_form.jsp}{{\tt Free Form SQL
Query}} page the Database release {\tt WSERV1000v20180716} can be
accessed which contains all the data we use here.

We {\it nota bene} a few things. First, the quantity {\tt aperJky3}
and {\tt aperJky3Err} are found in the {\tt wserv1000MapRemeasAver}
and {\tt wserv1000MapRemeasurement}, so care has to be taken to return
unique column names (otherwise e.g.
\href{http://docs.astropy.org/en/stable/io/fits/}{astropy.io.fits}
will crash).  As such, we alias {\tt aver.aperJky3} to {\tt
aperJky3Aver} and likewise for the error quantity. Aliases will be
necessary in some cases anyway, because some queries can be done
sensibly on multiple instances of the same table. Other times, one may
join tables on quantities such as {\tt catalogueID} or {\tt
apertureID}, where you are meaning the same thing, but aliases would
again be sensible.

Second, the {\tt RA} and {\tt DEC} values returned by the WSA are in radians, if
used directly. To return values in degrees, use a selection with an alias, e.g. 
{\tt RA as RADeg} and {\tt DEC as DECDeg}.

@@

At the \href{http://horus.roe.ac.uk/vsa/login.html}{VSA Login}, enter 
with these credentials::
\begin{itemize}
    \item Username: {\tt VSERV1000} 
    \item password: {\tt highzqso} 
    \item community: {\tt proprietary}
\end{itemize}
Then head to the \href{http://horus.roe.ac.uk:8080/vdfs/VSQL_form.jsp}{Freeform SQL Query} page where the database release to use is {\tt VSERV1000v20180716}.

>>>>

\title[Infrared properties Very high-$z$ Quasars]
  {Near and Mid-infrared properties of known $z\geq5$ Quasars}
\author[Ross \& Cross]
%\author[N. Ross, N. Cross \& C. Ross]
{Nicholas P. Ross\thanks{E-mail: npross@roe.ac.uk} and Nicholas J. G. Cross \\ 
%{Nicholas P. Ross\thanks{E-mail: npross@roe.ac.uk}, Nicholas J. G. Cross and Christine B. Ross \\ 
Institute for Astronomy, University of Edinburgh, Royal Observatory, Edinburgh, EH9 3HJ, United Kingdom\\ }

@@

Using the extended datasets described in Section~\ref{sec:NIR_data},
we select objects with at least 8 measurements observed with intervals
of at least 30 days, in at least one filter. The the average
calibrated flux over all epochs must be $>0$ (i.e. {\tt aperJky3}$>0$)
with the signal-to-noise is required to be $\geq8$. (i.e. {\tt
aperJky3/aperJky3Err}$\geq$8). The clipped median and median absolute
deviation is then calculated
\begin{equation}
      var = 1.48 \times  {\rm m.a.d.} / \bar{\epsilon}
\end{equation}
where $var$ is the index of variation, m.a.d. is the median absolute
deviation, and $\bar{\epsilon}$ is the mean of the error in each point
in the light curve, divided by the total number of points.  With these
criteria, there are 21 WFCAM and 12 VISTA objects with 1 object (VIK
J1148+0056) in common to both, and this sample of 32 is our starting
point for variability investigations.

@@

\begin{figure}
  \begin{subfigure}{}
    \centering
    \includegraphics[width=8.5cm]{../light_curves/MMTJ0215-0529LC_20190214.png}
    \caption{The infrared light curve for MMT J0215-0529 with data from 
      WFCAM (smaller solid circles), VIRCAM (hexagons) and NEOWISE-R (larger solid circles).}
    \label{fig:MMTJ0215-0529}
  \end{subfigure} 
  \begin{subfigure}{}
    \centering
    \includegraphics[width=8.5cm]{../light_curves/CFHQSJ0216-0455LC_20190214.png}
    \caption{The same as Fig.~\ref{fig:MMTJ0215-0529} except for CFHQS J0216-0455.}
    \label{fig:CFHQSJ0216-0455}
  \end{subfigure}
\end{figure}

@@

Even with well sampled data across the 3000-4000 day observed
time-scales that the WFCAM and VISTA surveys span, the $(1+z)$
time-dilation dramatically affects the sampled rest-frame timescales
sampled, which are $\sim300-700$ days. Indeed, when sharp changes in a accretion rate are expected
on the system's dynamical timescale of several kiloyears
\citep[e.g.,][]{Regan2019}, then seeing any variability signature is
not expected. However, noting how in lower redshift quasars, which
also have massive $>10^{8}$ M$_{\odot}$ black holes, dramatic changes
in both continuum and line emission {\it is seen} on much shorter
timescales, continued monitoring of these objects is warranted.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%  S E C T I O  N   7         S E C T I O  N   7           S E C T I O  N   7       S E C T I O  N   7
%%  S E C T I O  N   7         S E C T I O  N   7           S E C T I O  N   7       S E C T I O  N   7
%%  S E C T I O  N   7         S E C T I O  N   7           S E C T I O  N   7       S E C T I O  N   7
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}\label{sec:conclusions}
In this study, we have, for the first time, compiled the list of all
$z>5$ spectroscopically confirmed quasars. We have assembled the NIR
($y/Y, J, H, K/K_{s}$) and MIR (WISE W1/2/3/4) photometry for these
objects, given their detection rates and SEDs. We find that:

\begin{itemize}
    \item SDSS and Pan-STARRS1 together identified over half of the VH$z$Q sample;  
    \item There remains a quasar ``redshift desert'' at $z\approx5.3-5.7$, though 
      efforts are being made to address this \citep[e.g.,][]{YangJ2018a};
    \item 97.0\% of the VH$z$Q sample is detected in one or more NIR ($ZYJHK/K_{s}$) band; 
    \item The 14 objects that are not detected in the NIR are due to lack of coverage rather than lack of depth; 
   \item 362 ( 78.2\%) VH$z$Qs are detected by WISE, e.g. in the deeper unWISE W1 catalog. 
 \item All of the $z\geq7$ quasars are detected in both unWISE W1 and W2. 
 \item 32 of the QSOs had enough NIR measurements and a good enough signal-to-noise ratio to look for variability. Weak variability was detected in multiple bands of SDSS J0959+0227, and very marginally in the Y-band of MMT J0215-0529. 2 other QSOs, SDSS J0349+0034 and SDSS J2220-0101 had significant differences between their WFCAM and VISTA magnitudes in one band, also indicating variability. 
\end{itemize}

@@

The VISTA Data Flow System pipeline processing and science archive were used for the WFCAM and VISTA 
near infrared data are described in Irwin et al (2004), Hambly et al (2008) and Cross et al. (2012).

@@

%% N.P.R. may have listened to Kris Kross during the preparation of this manuscript.

@@

\section{Near-Infrared WFCAM Science Archive SQL queries}\label{sec:SQL}
Here we give the recipe and SQL that returned the near-infrared photometry 
for the VH$z$Qs from the  WFCAM Science Archive.

The data are on the WFCAM Science Archive: \href{wsa.roe.ac.uk}{\tt wsa.roe.ac.uk}. 
Access the User Login form \href{WFCAM Science Archive}{\tt wsa.roe.ac.uk/login.html} 
with these credentials::
\begin{itemize}
    \item Username: {\tt WSERV1000} 
    \item password: {\tt highzqso} 
    \item community: {\tt nonsurvey}
\end{itemize}
Then going to the
\href{http://wsa.roe.ac.uk:8080/wsa/SQL_form.jsp}{{\tt Free Form SQL
Query}} page the Database release {\tt WSERV1000v20190507} can be
accessed which contains all the data we use here.

We {\it nota bene} a few things. First, the quantity {\tt aperJky3}
and {\tt aperJky3Err} are found in the {\tt wserv1000MapRemeasAver}
and {\tt wserv1000MapRemeasurement}, so care has to be taken to return
unique column names (otherwise e.g.
\href{http://docs.astropy.org/en/stable/io/fits/}{astropy.io.fits}
will crash).  As such, we alias {\tt aver.aperJky3} to {\tt
aperJky3Aver} and likewise for the error quantity. Aliases will be
necessary in some cases anyway, because some queries can be done
sensibly on multiple instances of the same table. Other times, one may
join tables on quantities such as {\tt catalogueID} or {\tt
apertureID}, where you are meaning the same thing, but aliases are then necessary 
for SQLServer to correctly comprehend the query.

Second, the {\tt RA} and {\tt DEC} values returned by the WSA are in radians, if
used directly. To return values in degrees, use a selection with an alias, e.g. 
{\tt RA as RADeg} and {\tt DEC as DECDeg}. Other relevant query examples can be found in the
GitHub repository, under WSA_VSA/SAMPLE_SQL_QUERIES, or the SQLCookbook in the WSA or VSA.

@@

At the \href{http://horus.roe.ac.uk/vsa/login.html}{VSA Login}, enter 
with these credentials::
\begin{itemize}
    \item Username: {\tt VSERV1000} 
    \item password: {\tt highzqso} 
    \item community: {\tt proprietary}
\end{itemize}
Then head to the \href{http://horus.roe.ac.uk:8080/vdfs/VSQL_form.jsp}{Freeform SQL Query} page where the database release to use is {\tt VSERV1000v20190508}. The queries are the same as those used in the WSA, see Appendix~\ref{sec:SQL}, but the names of the tables starting with {\tt wserv1000} should be replaced with {\tt vserv1000}.
